---
title: "The distribution of sample means obtained from an exponential distribution"
author: "Jae Wook Lee"
date: "October 23, 2015"
output: html_document
---

##Overview

In this document, the properties of sample means  obtained from an exponential distribution are analyzed in terms of their relationship to the population mean and variance.

##Simulations
The following R code simulates sampling of 1000 sets of 40 exponentials. The rate parameter for the exponential distribution (lambda) is set to 0.2.  

```{r}
lambda = 0.2   # rate for the exponential distribution
nosim=1000     # number of simulations
n = 40         # size of a sample
samples = replicate(nosim, mean(rexp(n, rate=lambda)))  # a vector of 1000 sample means
```

##Sample mean versus Theoretical Mean

The mean of the sample means are `r mean(samples)`. It is almost equal to the population mean `r 1/lambda`.
In the following code, we can see that the mean of the sample means approaches the theoretical mean as the number of simulation grows.

```{r, echo=TRUE, fig.width=4, fig.height=3}
means <- sapply(1:nosim, function(x) mean(replicate(x, mean(rexp(n, lambda)))))
require(ggplot2)
g <- ggplot(data.frame(x = 1 : nosim, y = means), aes(x = x, y = y)) + geom_line(size=1.5)
g <- g + geom_hline(yintercept = 1/lambda, col="red", size=2)
g <- g + labs(x = "Number of simulation", y = "Mean of the sample means")
g
```

##Sample variance vs Theoretical variance
The theoretical variance for the exponential distribution with rate = 0.2 is (1/0.2)^2 = 25. The variance of 1000 sample means is `r var(samples)`, which is very close to the population variance divided by the size of the sample (40) = 25/40 = 0.625. Similar to the mean of the sample means, the variance of the sample means approaches population variance/sqrt(40) as the number of simulation grows.
```{r, echo=FALSE, fig.width=4, fig.height=3}
vars <- sapply(2:nosim, function(x) var(replicate(x, mean(rexp(n, lambda)))))
g <- ggplot(data.frame(x = 2 : nosim, y = vars), aes(x = x, y = y)) + geom_line(size=1.5)
g <- g + geom_hline(yintercept = (1/lambda)^2/n, col="red", size=2)
g <- g + labs(x = "Number of simulation", y = "Variance of the sample means")
g
```


##The distribution of 1000 sample means is approximately normal.

The histogram for the sample means shows that they approximate a normal distribution with a mean of 1/lambda and standard deviation of (1/lambda)/(sqrt(n)).

```{r, echo=FALSE, fig.width=4, fig.height=3}

p0 = qplot(samples, geom = 'blank') + 
        geom_histogram(aes(y = ..density..), color="black", fill="white") +
        stat_function(fun = dnorm, args=list(mean=1/lambda, sd=(1/lambda)/sqrt(n)), geom="line")+    
        theme(legend.position="none") +
        labs(title="Distribution of 1000 sample means", x="Mean of 40 exponentials", y="Density")
p0
```
  
  
   The quantile-quantile plot below also shows that 1000 samples means are normally distributed.
```{r, echo=FALSE, fig.width=4, fig.height=3}
qqnorm(samples)
```